---
title: "main"
output: html_document
---

Plan:
- Predict released movie revenue
- Predictors:
  - adult
  - popularity
  - budget
  - release date -> season
  - runtime
  - genres
  - casts (and their popularity?)
  - non-casts (and their popularity?)
  
```{r setup, eval = FALSE}
!install.packages(c("httr", "dplyr", "jsonlite", "progress", "dplyr", "tidyr", "ggplot2", "knitr", "ggcorrplot", "stringr", "xgboost"))
```

```{r library}
library(httr)
library(dplyr)
library(jsonlite)
library(progress)
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(ggcorrplot)
library(stringr)
library(xgboost)
```

```{r constants}
API_KEY <- "e6d34aa453d1a71af2b6e77848923e57"
BASE_URL <- "https://api.themoviedb.org/3/"
DISCOVER_MOVIE_PATH <- "discover/movie"
MOVIE_PATH <- "movie/"

MOVIES_COUNT <- 1000
MOVIES_PER_PAGE <- 20
```

# Fetching Data
```{r}
response <- GET(
  url = paste0(BASE_URL, DISCOVER_MOVIE_PATH),
  query = list(
    sort_by = "revenue.desc",
    api_key = API_KEY
  )
)

movie_total_results <- content(response)$total_results
movie_total_pages <- content(response)$total_pages

cat("There are", movie_total_results, "movies in total provided by the API.\n")
cat("These movies are paginated into", movie_total_pages, "pages", "with", MOVIES_PER_PAGE, "movies per page.\n")
```
## Fetch movie ids and titles
```{r}
movies <- data.frame()
n_pages <- ceiling(MOVIES_COUNT / MOVIES_PER_PAGE)
pb <- txtProgressBar(min = 0, max = n_pages, style = 3)

for (page in 1:n_pages) {
  response <- GET(
    url = paste0(BASE_URL, DISCOVER_MOVIE_PATH),
    query = list(
      sort_by = "revenue.desc",
      page = page,
      api_key = API_KEY
    )
  )
  
  # Check if API request was successful
  if (status_code(response) != 200) {
    print(paste("Skipping page", page, "- API error:", status_code(response)))
    next
  }
  
  data <- fromJSON(content(response, as = "text"))$results
  page_movies <- data.frame(
    movie_id = data$id,
    title = data$title
  )
  movies <- rbind(movies, page_movies)
  setTxtProgressBar(pb, page)
}

close(pb)
cat("Successfully fetch all movie ids and titles!")
```

## Fetch movie details (revenue, adult, popularity, budget, release date, runtime, genres)
```{r}
pb <- txtProgressBar(min = 0, max = nrow(movies), style = 3)
all_genres <- c()

for (i in 1:nrow(movies)) {
  movie_id <- movies$movie_id[i]
  
  response <- GET(
    url = paste0(BASE_URL, MOVIE_PATH, movie_id),
    query = list(
      api_key = API_KEY
    )
  )
  
  # Check if API request was successful
  if (status_code(response) != 200) {
    print(paste("Skipping movie", movie_id, "- API error:", status_code(response)))
    next
  }
  
  data <- fromJSON(content(response, as = "text"))

  movies$revenue[i] <- data$revenue
  movies$adult[i] <- as.logical(data$adult)
  movies$popularity[i] <- data$popularity
  movies$budget[i] <- data$budget
  movies$release_date[i] <- data$release_date
  movies$runtime[i] <- data$runtime

  genres <- data$genres
  movies$genres[i] <- paste(genres$name, collapse = ",")
  for (genre in genres$name) {
    all_genres <- unique(c(all_genres, genre))
  }
  setTxtProgressBar(pb, i)
}

close(pb)
cat("Successfully fetch movie details!")
```

```{r}
all_genres
```

## Fetch movie credits (casts and non-casts)
```{r}
cat(paste0(BASE_URL, MOVIE_PATH, 1231, "credits"))
pb <- txtProgressBar(min = 0, max = nrow(movies), style = 3)

movies$cast <- NA       # Stores top 5 actors
movies$non_cast <- NA   # Stores top 5 crew members (excluding actors)

for (i in 1:nrow(movies)) {
  movie_id <- movies$movie_id[i]
  
  response <- GET(
    url = paste0(BASE_URL, MOVIE_PATH, movie_id, "/credits"),
    query = list(
      api_key = API_KEY 
    )
  )
  
  # Check if API request was successful
  if (status_code(response) != 200) {
    print(paste("Skipping movie", movie_id, "- API error:", status_code(response)))
    next
  }

  credits <- fromJSON(content(response, as = "text"))

  # Extract CAST (Actors)
  if (!is.null(credits$cast) && length(credits$cast) > 0) {
    # Filter only actors (those with a "character" field)
    cast_members <- credits$cast[!is.null(credits$cast$character), ]
    
    top_cast <- head(cast_members, 5)
    
    # Format: "Actor1 (popularity), Actor2 (popularity)"
    movies$cast[i] <- paste0(top_cast$name, "(", round(top_cast$popularity, 1), ")", collapse = ", ")
    movies$actor_avg_pop[i] <- mean(top_cast$popularity)
  } else {
    movies$cast[i] <- NA  # Handle missing actor
    movies$actor_avg_pop[i] <- NA
  }
  
  # Extract NON-CAST (Directors, Producers, etc.)
  if (!is.null(credits$crew) && length(credits$crew) > 0) {
    # Exclude actors (crew members without "character" role)
    non_cast_members <- credits$crew[is.null(credits$crew$character), ]
    
    top_non_cast <- head(non_cast_members, 5)
    
    # Format: "Crew1 (popularity), Crew2 (popularity)"
    movies$non_cast[i] <- paste0(top_non_cast$name, "(", round(top_non_cast$popularity, 1), ")", collapse = ", ")
    movies$crew_avg_pop[i] <- mean(top_non_cast$popularity)
  } else {
    movies$non_cast[i] <- NA  # Handle missing crew
    movies$crew_avg_pop[i] <- NA
  }

  setTxtProgressBar(pb, i)
}
close(pb)
```
## Change column names
```{r}
colnames(movies) <- c(
  "movie_id",      # Unique movie ID
  "title",         # Movie title
  "revenue",       # Box office revenue
  "adult",         # Whether the movie is rated adult (TRUE/FALSE)
  "popularity",    # Popularity score from TMDB
  "budget",        # Production budget
  "release_date",  # Release date
  "runtime",       # Duration in minutes
  "genres",        # Movie genres
  "actors",        # Top 5 actors
  "crews",         # Top 5 crew members
  "actor_avg_pop", # Average popularity of the top 5 actors
  "crew_avg_pop"   # Average popularity of the top 5 crew members
)
colnames(movies)
```

# Cleaning the data

## Store a copy of the original data
```{r}
movies_original <- movies
```

## Convert `release_date` to date
```{r}
movies$release_date <- as.Date(movies$release_date)
```

## Check for missing values
```{r}
colSums(is.na(movies))
```

```{r}
cat("Summary for movie_id:\n")
summary(movies$movie_id)
cat("Number of empty title: ", sum(movies$title == ""), "\n")
cat("Summary for revenue:\n")
summary(movies$revenue)
cat("Unique values of adult:\n")
unique(movies$adult)
cat("Summary for popularity:\n")
summary(movies$popularity)
cat("Summary for budget:\n")
summary(movies$budget)
cat("Summary for release_date:\n")
summary(movies$release_date)
cat("Summary for runtime:\n")
summary(movies$runtime)
cat("Unique values of genres:\n")
unique(unique(unlist(strsplit(movies$genres, ","))))
```

We observe that all values are normal, except that adult is always false, some movies have 0 budget, some movies have 0 runtime, some movies don't have any actor or crew, and there are no release date for 4 movies.

## Adult is always false
Since adult is always false, we're not going to include it into our analysis since it doesn't provide any insight.
```{r}
movies <- select(movies, -adult)
```

## 0 budget movies
```{r}
zero_budget <- movies$budget == 0
movies[zero_budget,]
```

## 0 runtime movies
```{r}
zero_runtime <- movies$runtime == 0
movies[zero_runtime,]
```

## Remove zero values
```{r}
movies <- filter(movies, budget > 0 & runtime > 0)
```

## Movies without actor
```{r}
actors_na <- is.na(movies$actors)
movies[actors_na,]
```

## Movies without crew
```{r}
crews_na <- is.na(movies$crews)
movies[crews_na,]
```

## Movies without release date
```{r}
release_date_na <- is.na(movies$release_date)
movies[release_date_na,]
```

## Remove NA values
```{r}
movies <- filter(movies, !is.na(release_date) & !is.na(actors) & !is.na(crews))
colSums(is.na(movies))
```

##  Recheck the values
```{r}
cat("Summary for budget:\n")
summary(movies$budget)
cat("Summary for runtime:\n")
summary(movies$runtime)
```

## Remove movies with an out-of-place runtime
```{r}
movies <- filter(movies, runtime != 4)
summary(movies$runtime)
```

## Check the dimension, header, and footer of our cleaned dataset
```{r}
cat("Shape of movies: ", dim(movies), "\n")
head(movies)
tail(movies)
```

# Pre-processing data

## Convert `release_date` to `season`
```{r}
movies$season <- as.factor(case_when(
  format(movies$release_date, "%m") %in% c("12", "01", "02") ~ "Winter",
  format(movies$release_date, "%m") %in% c("03", "04", "05") ~ "Spring",
  format(movies$release_date, "%m") %in% c("06", "07", "08") ~ "Summer",
  format(movies$release_date, "%m") %in% c("09", "10", "11") ~ "Fall",
  TRUE ~ "Unknown"
))
levels(movies$season)
```

## Convert `genres` into one-hot encoding
```{r}
movies <- movies %>%
  separate_rows(genres, sep = ",") %>%
  mutate(value = 1) %>%
  pivot_wider(names_from = genres, values_from = value, values_fill = list(value = 0))
```

## Check the dimension, NA values, header, and footer of our cleaned dataset
```{r}
cat("Shape of movies: ", dim(movies), "\n")
cat("Total number of missing values across all columns =", sum(colSums(is.na(movies))), "\n")
head(movies)
tail(movies)
```

# Exploratory data analysis (EDA)

## Numerical summaries
```{r}
cat("Summary statistics for key numeric or date variables:\n")
summary(select(movies, revenue, popularity, budget, release_date, runtime, actor_avg_pop, crew_avg_pop))

cat("\nCount of movies for each season:")
table(movies$season)
```

## Response variable (revenue)
```{r}
# Original scale
ggplot(movies, aes(x = revenue)) +
  geom_histogram(bins = 30, fill = "#3498db", color = "white", alpha = 0.8) +
  theme_minimal() +
  labs(title = "Distribution of Movie Revenue", 
       x = "Revenue", 
       y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

# Log scale
ggplot(movies, aes(x = log10(revenue))) +
  geom_histogram(bins = 30, fill = "#3498db", color = "white", alpha = 0.8) +
  theme_minimal() +
  labs(title = "Distribution of Movie Revenue (Log Scale)", 
       x = "Log10(Revenue)", 
       y = "Frequency") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
```

## Revenue vs. Budget
```{r}
ggplot(movies, aes(x = budget, y = revenue)) +
  geom_point(alpha = 0.4, color = "#e74c3c") +
  geom_smooth(method = "lm", color = "#2c3e50", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Budget vs. Revenue", 
       x = "Budget", 
       y = "Revenue") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

ggplot(movies, aes(x = log10(budget), y = log10(revenue))) +
  geom_point(alpha = 0.4, color = "#e74c3c") +
  geom_smooth(method = "lm", color = "#2c3e50", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Budget vs. Revenue", 
       x = "Log10(Budget)", 
       y = "Log10(Revenue)") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
```

## Revenue vs. Popularity
```{r}
## Revenue vs. Popularity
ggplot(movies, aes(x = popularity, y = revenue)) +
  geom_point(alpha = 0.5, color = "#3498db") +
  geom_smooth(method = "lm", color = "#2c3e50", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Movie Popularity vs. Revenue",
    x = "TMDb Popularity Score",
    y = "Revenue"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13, face = "bold")
  )

ggplot(movies, aes(x = popularity, y = log10(revenue))) +
  geom_point(alpha = 0.5, color = "#3498db") +
  geom_smooth(method = "lm", color = "#2c3e50", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Movie Popularity vs. Revenue",
    x = "TMDb Popularity Score",
    y = "Log10(Revenue)"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13, face = "bold")
  )
```

## Revenue vs. Runtime
```{r}
ggplot(movies, aes(x = runtime, y = revenue)) +
  geom_point(alpha = 0.5, color = "#3498db") +
  geom_smooth(method = "lm", col = "#2c3e50", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Movie Runtime vs. Revenue",
       x = "Runtime (minutes)",
       y = "Revenue") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13, face = "bold")
)

ggplot(movies, aes(x = runtime, y = log10(revenue))) +
  geom_point(alpha = 0.5, color = "#3498db") +
  geom_smooth(method = "lm", col = "#2c3e50", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Movie Runtime vs. Revenue",
       x = "Runtime (minutes)",
       y = "Log10(Revenue)") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13, face = "bold")
)
```

## Revenue vs. Release Season
```{r}
ggplot(movies, aes(x = season, y = log10(revenue), fill = season)) +
  geom_boxplot(alpha = 0.85, outlier.shape = 21, outlier.fill = "white", outlier.color = "black") +
  scale_fill_manual(values = c("Winter" = "#74b9ff",
                               "Spring" = "#00b894",
                               "Summer" = "#f9ca24",
                               "Fall" = "#d35400")) +
  theme_minimal() +
  labs(title = "Movie Revenue by Release Season", 
       x = "Season", 
       y = "Log10(Revenue)") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(face = "bold", size = 12)
  )
```

## Revenue vs. Release Year
```{r}
library(lubridate)

# Extract year from release date
movies$release_year <- year(movies$release_date)

# Compute average revenue per year
yearly_revenue <- movies %>%
  group_by(release_year) %>%
  summarise(avg_revenue = mean(revenue, na.rm = TRUE), movie_count = n()) %>%
  arrange(release_year)

# Plot time series
ggplot(yearly_revenue, aes(x = release_year, y = avg_revenue)) +
  geom_line(color = "#e74c3c", size = 1.5) +  # Bold red line
  geom_point(color = "#2c3e50", size = 3) +  # Dark gray dots
  theme_minimal() +
  labs(
    title = "Average Movie Revenue Over Time",
    x = "Release Year",
    y = "Average Revenue ($)"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +  # Format revenue as dollars
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "#2c3e50"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13, face = "bold")
  )
```

## Revenue vs. Genres
```{r}
# Define colors based on genre themes
genre_colors <- c(
  "Action" = "#e74c3c",      
  "Adventure" = "#f39c12",   
  "Animation" = "#f1c40f",      
  "Comedy" = "#2ecc71",       
  "Crime" = "#34495e",        
  "Documentary" = "#95a5a6",    
  "Drama" = "#9b59b6",     
  "Family" = "#e67e22",    
  "Fantasy" = "#8e44ad",       
  "History" = "#d35400",       
  "Horror" = "#c0392b",     
  "Music" = "#1abc9c",     
  "Mystery" = "#16a085",       
  "Romance" = "#e84393",      
  "Science Fiction" = "#27ae60",
  "TV Movie" = "#7f8c8d",      
  "Thriller" = "#2c3e50",     
  "War" = "#bdc3c7",            
  "Western" = "#8d6e63"
)

# Prepare genre-specific revenue data
genre_revenue <- movies %>%
  select(starts_with("Action"):starts_with("Documentary"), revenue) %>%
  pivot_longer(cols = starts_with("Action"):starts_with("Documentary"), names_to = "Genre", values_to = "Presence") %>%
  filter(Presence == 1)

# Boxplot of revenue by genre
ggplot(genre_revenue, aes(x = reorder(Genre, log10(revenue), FUN = median), y = log10(revenue), fill = Genre)) +
  geom_boxplot(alpha = 0.85, outlier.shape = 21, outlier.fill = "white", outlier.color = "black") +
  scale_fill_manual(values = genre_colors) +
  coord_flip() +  # Horizontal layout for readability
  theme_minimal() +
  labs(title = "Revenue Distribution by Genre",
       x = "Genre",
       y = "Log10(Revenue)") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "none",
    axis.text.x = element_text(face = "bold", size = 12),
    axis.text.y = element_text(face = "bold", size = 10)
  )

```

## Revenue vs. Actor's popularity
```{r}
ggplot(movies, aes(x = actor_avg_pop, y = log10(revenue))) +
  geom_point(alpha = 0.5, color = "#9b59b6") +
  geom_smooth(method = "lm", col = "#2c3e50", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Actor Popularity vs. Revenue", 
       x = "Average Actor Popularity", 
       y = "Log10(Revenue)") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
```

## Revenue vs. Crew's popularity
```{r}
ggplot(movies, aes(x = crew_avg_pop, y = log10(revenue))) +
  geom_point(alpha = 0.5, color = "#e74c3c") +
  geom_smooth(method = "lm", col = "#2c3e50", se = FALSE, linetype = "dashed") +
  theme_minimal() +
  labs(title = "Crew Popularity vs. Revenue",
       x = "Average Crew Popularity",
       y = "Log10(Revenue)") +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13, face = "bold")
  )
```

## Top actors with highest average revenue, total revenue, and movie count
```{r}
# Extract actor names without popularity score
top_actors_total <- movies %>%
  separate_rows(actors, sep = ", ") %>%
  mutate(actor_name = str_extract(actors, "^[^(]+")) %>%  # Extract only the name, ignoring numbers
  group_by(actor_name) %>%
  summarise(total_revenue = sum(revenue, na.rm = TRUE)) %>%
  arrange(desc(total_revenue)) %>%
  slice_max(total_revenue, n = 20, with_ties = FALSE)  

# Plot
ggplot(top_actors_total, aes(x = reorder(actor_name, total_revenue), y = total_revenue, fill = total_revenue)) +
  geom_bar(stat = "identity", alpha = 0.9) +
  coord_flip() +
  scale_fill_gradient(low = "#F5B7B1", high = "#922B21") +  # Light Pink → Deep Red
  theme_minimal() +
  labs(
    title = "Top 20 Actors by Total Movie Revenue",
    x = "Actor",
    y = "Total Revenue ($)",
    fill = "Total Revenue"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +  
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "#641E16"),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12, face = "bold")
  )


# Extract top actors by average revenue per movie
top_actors_avg <- movies %>%
  separate_rows(actors, sep = ", ") %>%
  mutate(actor_name = str_extract(actors, "^[^(]+")) %>%  
  group_by(actor_name) %>%
  summarise(avg_revenue = mean(revenue, na.rm = TRUE)) %>%
  arrange(desc(avg_revenue)) %>%
  slice_max(avg_revenue, n = 20, with_ties = FALSE)  

# Plot
ggplot(top_actors_avg, aes(x = reorder(actor_name, avg_revenue), y = avg_revenue, fill = avg_revenue)) +
  geom_bar(stat = "identity", alpha = 0.9) +
  coord_flip() +
  scale_fill_gradient(low = "#F8DE7E", high = "#D4AC0D") +  # Light Gold → Deep Gold
  theme_minimal() +
  labs(
    title = "Top 20 Actors by Average Movie Revenue",
    x = "Actor",
    y = "Average Revenue per Movie ($)",
    fill = "Avg Revenue"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +  
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "#7D6608"),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12, face = "bold")
  )


# Extract top actors by movie count
top_actor_counts <- movies %>%
  separate_rows(actors, sep = ", ") %>%
  mutate(actor_name = str_extract(actors, "^[^(]+")) %>%  
  group_by(actor_name) %>%
  summarise(movie_count = n()) %>%
  arrange(desc(movie_count)) %>%
  slice_max(movie_count, n = 20, with_ties = FALSE)  

# Plot
ggplot(top_actor_counts, aes(x = reorder(actor_name, movie_count), y = movie_count, fill = movie_count)) +
  geom_bar(stat = "identity", alpha = 0.9) +
  coord_flip() +
  scale_fill_gradient(low = "#FAD7A0", high = "#D35400") +  # Light Orange → Deep Orange
  theme_minimal() +
  labs(
    title = "Top 20 Actors by Number of Movies",
    x = "Actor",
    y = "Movie Count",
    fill = "Movie Count"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "#B34700"),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12, face = "bold")
  )
```

## Top crews with highest average revenue, total revenue, and movie count
```{r}
# Extract top crew members by total revenue
top_crew_total <- movies %>%
  separate_rows(crews, sep = ", ") %>%
  mutate(crew_name = str_extract(crews, "^[^(]+")) %>%  
  group_by(crew_name) %>%
  summarise(total_revenue = sum(revenue, na.rm = TRUE)) %>%
  arrange(desc(total_revenue)) %>%
  slice_max(total_revenue, n = 20, with_ties = FALSE)  

# Plot
ggplot(top_crew_total, aes(x = reorder(crew_name, total_revenue), y = total_revenue, fill = total_revenue)) +
  geom_bar(stat = "identity", alpha = 0.9) +
  coord_flip() +
  scale_fill_gradient(low = "#A2D9CE", high = "#117A65") +  # Light Teal → Deep Green
  theme_minimal() +
  labs(
    title = "Top 20 Crew Members by Total Movie Revenue",
    x = "Crew Member",
    y = "Total Revenue ($)",
    fill = "Total Revenue"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +  
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "#154360"),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12, face = "bold")
  )


# Extract top crew members by average revenue per movie
top_crew_avg <- movies %>%
  separate_rows(crews, sep = ", ") %>%
  mutate(crew_name = str_extract(crews, "^[^(]+")) %>%  
  group_by(crew_name) %>%
  summarise(avg_revenue = mean(revenue, na.rm = TRUE)) %>%
  arrange(desc(avg_revenue)) %>%
  slice_max(avg_revenue, n = 20, with_ties = FALSE)  

# Plot
ggplot(top_crew_avg, aes(x = reorder(crew_name, avg_revenue), y = avg_revenue, fill = avg_revenue)) +
  geom_bar(stat = "identity", alpha = 0.9) +
  coord_flip() +
  scale_fill_gradient(low = "#85C1E9", high = "#1B4F72") +  # Light Blue → Deep Blue
  theme_minimal() +
  labs(
    title = "Top 20 Crew Members by Average Movie Revenue",
    x = "Crew Member",
    y = "Average Revenue per Movie ($)",
    fill = "Avg Revenue"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +  
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "#154360"),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12, face = "bold")
  )


# Extract top crew members by movie count
top_crew_counts <- movies %>%
  separate_rows(crews, sep = ", ") %>%
  mutate(crew_name = str_extract(crews, "^[^(]+")) %>%  
  group_by(crew_name) %>%
  summarise(movie_count = n()) %>%
  arrange(desc(movie_count)) %>%
  slice_max(movie_count, n = 20, with_ties = FALSE)  

# Plot
ggplot(top_crew_counts, aes(x = reorder(crew_name, movie_count), y = movie_count, fill = movie_count)) +
  geom_bar(stat = "identity", alpha = 0.9) +
  coord_flip() +
  scale_fill_gradient(low = "#D7BDE2", high = "#6C3483") +  # Light Purple → Deep Violet
  theme_minimal() +
  labs(
    title = "Top 20 Crew Members by Number of Movies",
    x = "Crew Member",
    y = "Movie Count",
    fill = "Movie Count"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "#512E5F"),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 12),
    axis.text.y = element_text(size = 12, face = "bold")
  )
```

## Correlation heatmap of key numeric variables
```{r}
# Select only numerical columns
numeric_vars <- select(movies, revenue, popularity, budget, runtime, actor_avg_pop, crew_avg_pop)

# Compute correlation matrix
corr_matrix <- cor(numeric_vars, use = "complete.obs")

# Define a custom theme for better aesthetics
ggcorrplot(corr_matrix, method = "square",
           colors = c("#1f77b4", "white", "#d62728"),  # Blue (negative) → White (neutral) → Red (positive)
           lab = TRUE,  # Show correlation values
           lab_size = 5,  # Larger labels
           outline.color = "black",  # Borders for better contrast
           ggtheme = theme_minimal()) +  # Minimalist theme
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "#2c3e50"),  # Elegant title
    panel.grid.major = element_blank(),  # Remove grid lines for a clean look
    panel.border = element_blank(),  # No outer border
    axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = "#2c3e50", face = "bold"),  # Rotated x-labels
    axis.text.y = element_text(size = 12, color = "#2c3e50", face = "bold")  # Bold y-labels
  ) +
  labs(title = "Correlation Heatmap: Movie Revenue vs Predictors")
```

# Some statistical analysis

## Linear relationship analysis for numeric predictors
```{r}
lm_model <- lm(log10(revenue) ~ log10(budget) + popularity + runtime + actor_avg_pop + crew_avg_pop, data = movies)

summary(lm_model)
```

## Linear relationship analysis for each genre
```{r}
lm_model <- lm(
  log10(revenue) ~ Action + Adventure + Fantasy + `Science Fiction` + Drama + Romance + 
    Animation + Comedy + Family + Thriller + Crime + History + Music + War + 
    Mystery + Horror + Western + Documentary, 
  data = movies
)

summary(lm_model)
```
```{r}
anova_model <- aov(log10(revenue) ~ season, data = movies)
summary(anova_model)
```
